---
title: "h5SE -- experiments with delayed HDF5 export/import for SummarizedExperiments"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{h5DSE: Out-of-memory illustrations for SummarizedExperiment}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

We want to get some background on the use of HDF5 and DelayedArray
infrastructure for modest-size assay sets.

We'll work with some 450k data.
```{r setup}
suppressPackageStartupMessages({
library(SummarizedExperiment)
library(HDF5Array)
library(yriMulti)
if (!exists("banovichSE")) data(banovichSE)
})
banovichSE
```

# Export a standard RangedSummarizedExperiment

Let's use the newly defined method for exporting the
assay component of a SummarizedExperiment:
```{r donew}
library(benchOOM)
tf = tempfile()
dir.create(tf)
banoSE = saveHDF5SummarizedExperiment(banovichSE, tf, replace=TRUE)
banoSE
```

Access capabilities:
```{r usenew}
assay(banoSE[1,])
subsetByOverlaps(banoSE, rowRanges(banoSE)[1001:1010])
```

# Import a packaged HDF5RangedSummarizedExperiment

```{r lkmm}
litg = getLitGeu()
litg
```

# Benchmarking exercise

We want to create a spectrum of biologically meaningful queries.  To
this end we introduced the gobrowse and go2ranges functions.  
```{r bb}
library(Organism.dplyr)
if (file.exists("~/hg19.sqlite")) 
  hg19 = src_organism(dbpath="~/hg19.sqlite") else {
  library(TxDb.Hsapiens.UCSC.hg19.knownGene)
  hg19 = src_organism(TxDb.Hsapiens.UCSC.hg19.knownGene)
  }
r1 = go2ranges("transmembrane receptor activity", hg19)
length(r1)
library(microbenchmark)
microbenchmark(subsetByOverlaps(banovichSE, r1), times=5)
microbenchmark(subsetByOverlaps(banoSE, r1), times=5)
```

